---
output:
  word_document: default
  html_document: default
---

# **Re-structuring bio-database**

## *Mutating a new bio-database on PostgreSQL* 

The bio-database of ODB had been constructed since 2009 and completely set-up on Microsoft SQL Server since 2015. The data curation is basically done by a Microsoft C# program. These protocols of data curation was documented in 圖輯. For developing an open cross-platform framework for web APIs of bio-database, first, we mutated a bio-database on PostgreSQL, i.e., exported from SQL Server, and then re-imported to a new database by using PostgreSQL.

<br>

::: {.infobox data-latex=""}

:::: {.box_main custom-style="box1_main"}
Box2.1
::::

:::: {.box_submain custom-style="box1_sect"}
Why not just use original Microsoft SQL Server as backend database for developing APIs of bio-database?
::::

:::: {custom-style="box1_inner"}
  &check; SQL Server is a commercial software with annual license fees.

  &check; SQL Server is suited for the Microsoft Server based framework, but not for an open cross-platform framework.  
::::

:::: {.box_submain custom-style="box1_sect"}
Why choose PostgreSQL?
::::

:::: {custom-style="box1"}
  &check; PostgreSQL is a powerful open-source object-relational database systems.

  &check; PostgreSQL can be well integrated in Geographic Information System (GIS) applications by its PostGIS extension.
  - It’s an important reason why we choose PostgreSQL. Most of APIs in ODB are used to develop GIS related applications.
::::

:::

<br>

As a backend database for API in production, we do not need some columns of variables originally in SQL Server (Fig. 2a) that used to remark some quality control (QC) issues. So when mutating the bio-database on PostgreSQL, we also did some simplifications for database schema (Fig. 2b). Roughly speaking, we left the “remark_xxx” or “flag_xxx” columns in SQL Server only, and in PostgreSQL, re-arranged the tables to make frequently-used variables more easily be accessed in table “cast_site” (site information) and “taxa_data” (abundance data of taxonomic groups). Another important modification is to construct classification of taxonomic groups in table ‘taxon_group” that can help us in analytical statistics among taxonomic groups when using bio-database data. The major parts of schema for bio-database on PostgreSQL include:
taxon_group: 

1. cast_site:

2. c00_cast_tbl:

3. t00_taxa_tbl:

4. taxa_data:

``` {r sqlServer2postgres, message=FALSE, warning=FALSE}
# NOW on Ubuntu, this driver not support old SQL Server 2008-R2. It only works on Windows
# On Ubuntu, the driver setting is in /etc/odbcinst.ini, e.g. driver={ODBC Driver 18 for SQL Server}
library(RODBC)      #R interface of ODBC driver to connect SQL Server
library(sqldf)      #runing SQL statements on R data frames
library(data.table) #data.table is superb fast in R data manipulation
library(magrittr)   #pipe function
ms_conn <- 
  paste0('driver={SQL Server};server=', sqlServerHost,
         ';database=', sqlServerDB,
         ';uid=', sqlServerUser,
         ';pwd=', sqlServerPass) %>%
  odbcDriverConnect()

# Query all taxonomic abundance in SQL server
taxa_data <- sqlQuery(ms_conn, 'select * from dbo.Taxa_record') %>% setDT()

# Query site
cast_site <- sqlQuery(ms_conn, 'select * from dbo.Cast') %>% setDT()

close(ms_conn) # Close database connection

```
::: {custom-style="Table Caption"}
Table `r run_bookmark("taxa_data", tab_zz)` Taxaonomic abundance in bio-database
:::

```{r taxa_data, echo=FALSE, tab.id='taxa_data', label='taxa_data'}
head(taxa_data[, .(taxarec_id, cast_id, taxonomic_name, taxon_count, original_unit)])

```

::: {custom-style="Table Caption"}
Table `r run_bookmark("cast_site", tab_zz)` Casting sites in bio-database
:::

```{r cast_site, echo=FALSE, tab.id='cast_site', label='cast_site'}
head(cast_site[, .(cast_id, station_id, date, depth_lower_bound, mesh_size, gear_type)])

```

``` {r updateTaxonGroup, message=FALSE, warning=FALSE}
library(RPostgres) #R (using Rcpp) Interface to PostgreSQL
library(odbapi) #Internal API package, which is explained in 
#e.g., A dataset with species Creseis conica, Abudefduf saxatilis, Labidocera gallensis, Metridia macrura, and M. asymmetrica is ready to be imported.
test_sp <- c("Creseis conica", "Abudefduf saxatilis", 
             "Labidocera gallensis", "Metridia macrura", "Metridia asymmetrica")

pconn <- dbConnect(drv = RPostgres::Postgres(),
                   host = odbHost, port = odbPort, dbname = odbBioDB,
                   user = odbUser, password = odbPass)
#Got taxonomy of bio-database on PostgreSQL
taxonomy <- dbReadTable(pconn, name="taxon_group") %>% setDT()
#getSciName first retrieve query from existed taxonomy in "taxon_group" table.
#For those scientific names not in bio-database, retrieve them by web API of WORMS, GBIF, etc. Internally its magic is done through R package "taxize".
new_taxa <- odbapi::getSciName(dbuser = odbUser, dbhost = odbHost, dbname = odbBioDB,
                               taxon = test_sp[!test_sp %chin% taxonomy$show_name],
#Only the latter three species are new to “taxon_group” table. This check can be ignored.
                               source = c("worms", "gbif")) #see Table 2.3
```

::: {custom-style="Table Caption"}
Table `r run_bookmark("taxonomy", tab_zz)` Update table of taxonomy (parts) in bio-database
:::

```{r taxonomy, echo=FALSE, tab.id='taxonomy', label='taxonomy'}
head(new_taxa[, .(query, rank, taxon, class, order, family, genus, species)])

```


## *Collaboration with larval fish database*

``` {r larvalFishDistribution, message=FALSE, warning=FALSE}
fconn <- dbConnect(drv = RPostgres::Postgres(),
                   host = larvaHost, port = odbPort, dbname = larvaDB,
                   user = odbUser, password = odbPass)
# Similar schema to get the same tables from two different data source on PostgreSQL
# Advantage: to design an integrated API to access them is more easier
bio_site <- dbReadTable(pconn, name="cast_site") %>% setDT()
larva_site <- dbReadTable(fconn, name="cast_site") %>% setDT()

library(sf) # R package that handles geometries in GIS applications 
sites <- list(bio_site[, source:="ODB"], larva_site[, source:="CHIU"]) %>% 
  rbindlist(use.names = TRUE, fill = TRUE) %>% 
  .[, season := odbapi::datex_season(date)] %>% # Internal API: convert date to season
  .[, season := factor(season, levels = c(0,1,2,3),
                       labels=c("Spring","Summer","Autumn","Winter"))] %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)
bbox <- st_bbox(sites) # Bounding box, used in plotting Fig. 2.4
dbDisconnect(pconn)    # Close database connection
dbDisconnect(fconn) 

```


::: {.fbox data-latex="" custom-style="fig1"}

```{r fig-distrib, echo=FALSE, message=FALSE, warning=FALSE, fig.id='fig-distrib', label='fig-distrib'}
library(ggplot2)
library(rnaturalearth)
ggplot(data=sites) + 
    geom_sf(aes(color = season), alpha=0.5, size=0.8) +
    facet_grid(. ~ source) +
    geom_path(data = ne_coastline(scale = 'medium'), aes(x = long, y = lat, group = group), color = 'darkgray', size = .3) +
    coord_sf(xlim=c(bbox[1]-1.5,bbox[3]+1.5), ylim=c(bbox[2]-1.5, bbox[4]+1.5)) + 
    xlab(NULL) + ylab(NULL) + 
    scale_color_viridis_d() +
    guides(color = guide_legend(title="Season:", override.aes = list(size = 2))) + 
    theme_light() + 
    theme(legend.position = "bottom", 
          #legend.key.size = unit(2, "cm"),
          legend.box.background = element_blank(), 
          legend.background = element_rect(fill = "transparent", colour = "transparent"),
          strip.text.y = element_text(angle = 0),#,face = "italic"),
          axis.text.x  = element_text(family = "sans"),
          axis.title.x = element_text(family = "sans"),
          axis.title.y = element_text(family = "sans"),
          axis.text.y = element_text(family = "sans"),
    )
```
:::

::: {custom-style="Image Caption"}
Fig. `r run_bookmark("fig-distrib", fig_zz)` Distribution of sampling sites in ODB bio-database
:::


## *Taxonomic composition of bio-database*

ODB bio-database collected a total of 137,819 records of abundance data, mostly zooplankton and larval fish, at 6700 sampling sites in 475 cruises. Fig. 2.5 shows the taxonomic composition in bio-database with the number of records and proportions. The largest proportion larval fish data, mainly comes from Dr. Chiu’s research, sums up to a 61,165 records with 673 species, 500 genera, and 190 families (Table 2.4).


## *Eco-environment database in ODB*

This is a reference to table \@ref(tab:tab1).

::: {custom-style="Table Caption"}
Table `r run_bookmark("tab1", tab_zz)` Dataset demo
:::


```{r tab1, echo=FALSE, tab.id='tab1', label='tab1'}
fread("./env_database202203.csv") %>% .[]
```

``` {r getenvGlossary, message=FALSE, warning=FALSE}
odbapi::getEnv_glossary()
```

